Finding and controling biases in dark matter halo concentration estimates

### 2016-07-09 paper/ms.tex
minor updates in abstract

\begin{abstract}
We use bootstrapping to estimate the bias of concentration
estimates on N-body dark matter halos as a function of particle
number. 
We find that algorithms based on the maximum radial velocity
and radial particle binning have increaing biases with decreasing 
particle numbers.
These methods overestimate the concentration by 
$15\%-20\%$ for halos sampled with $200$ particles, and $7\%$-$10\%$
for halos sampled with $500$ particles. 
To control this bias level we propose a new algorithm based on the
integrated mass profile. 
The method uses the full particle information without any binning,
making it reliable in cases when low numerical resolution becomes a
limitation for other methods.
This method reduces the bias to $2\%-3\%$ for halos sampled with
$200$-$500$ particles.
To keep the biases of the velocity and density methods to the same
low levels as in the new method, only halos sampled
with at least $\sim 4000$ particles should be used.
We also show that the mass-concentration relationship could be
shallower than expected from numerical experiments once the biases 
of the different concentration measurements are taken into account.
From these results we believe that bootstrapping and the
concentration estimation based on the integrated mass are
valuable tools to probe the internal structure of dark matter halos
in numerical simulations.
\end{abstract}

>>>>

\begin{abstract}
We use bootstrapping to estimate the bias of concentration
estimates on N-body dark matter halos as a function of particle
number. 
We find that algorithms based on the maximum radial velocity
and radial particle binning have increaing biases with decreasing 
particle numbers.
These methods overestimate the concentration by 
$15\%-20\%$ for halos sampled with $200$ particles, and $7\%$-$10\%$
for halos sampled with $500$ particles. 
To control this bias level we propose a new algorithm to estimate halo
concentrations based on the integrated mass profile. 
The method uses the full particle information without any binning,
making it reliable in cases when low numerical resolution becomes a
limitation for other methods.
This method reduces the bias to $2\%-3\%$ for halos sampled with
$200$-$500$ particles.
To keep the biases of the velocity and density methods to the same
low levels as in the integrated mass method, we find that only halos sampled
with at least $\sim 4000$ particles should be used.
We also show that the mass-concentration relationship could be
shallower than expected from numerical experiments once the biases 
of the different concentration measurements are taken into account.
From these results we believe that bootstrapping and the
concentration estimation based on the integrated mass profile are
valuable tools to probe the internal structure of dark matter halos
in numerical simulations.
\end{abstract}
